---
title: "covid 19"
author: "Bethanee Olmos"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   

$H_0$: There is no relationship between sex and amount of people dying from covid.


$H_A$: There is a relationship between sex and amount of people dying from covid.

# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="percent Bar Chart")
```

The percentage of females is less than males. The percentage of females with the virus is more but the percentage of females dying is less. The percentage of males with the virus is less however the percentage of deaths is more. 

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

56% of those alive are females and 36% are dead, 44% of those alive are male and 64% of those are dead. 
## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds of dying if you are a male is 44/1411, The odds of dying if you are a female is 25/1831. 
The odds ratio is (44/1411)/(25/1831)
The odds ratio is 2.3 to 1

## Conclusions

The p value is .001 which means that the chances of getting results that we just obtained if the null hypothesis is true. Based on the chances of being so low we reject the null hypothesis and accept that there is a relationship between sex and amount of people dying due to covid and conclude that men die at a higher rate than men do. 